---
title: "Horse EDA"
author: "Jeremiah W"
date: "May 2021"
output: html_document
---

# Potential Feature Engineering

1. number of consecutive/near-consecutive wins in the top 4, top X, etc.
    * e.g. horse XYZ placed in the top 4 three times in a row
2. for the same track & distance, standard deviation of winning times for horses that placed in the top 4
3. measure 2 relative to the medan/median time of horses for that track & distance
4. weight carried = actual - declared weight

```{r include=FALSE}
pacman::p_load(
  data.table,
  tidyverse,
  lubridate,
  dlookr,
  DataExplorer,
  inspectdf,
  funModeling,
  tictoc,
  beepr,
  ggthemes
)

#https://www.kaggle.com/gdaley/hkracing
a = read_csv('races.csv') 
b = read_csv('runs.csv')

a %>% head
b %>% head

```

```{r include=FALSE}
#weekday col
#a$wday = a$date %>% wday(label = T, abbr = T)

#prize in USD (used exchange rate as of May 23, 2021)

#a$prize.usd = a$prize/7.76413

#unused cols, all NAs
a$sec_time7 = NULL
a$time7 = NULL
# for now, not interested in the payout features -- theoretically shouldn't influence results of the race
a = a %>% select(-contains('dividend'))
```

```{r}
a = a %>% 
  mutate_if(is_character, factor) %>% 
  select(colnames(a) %>% sort) %>% 
  select(
    where(is.Date),
    where(is.factor),
    where(is.numeric)
  ) %>% relocate(date)

b = b %>% 
  mutate_if(is_character, factor) %>% 
  select(colnames(b) %>% sort) %>% 
  select(
    where(is.Date),
    where(is.factor),
    where(is.numeric)
  )

a %>% head
b %>% head
```
```{r}
a %>% filter(race_id %in% c(3)) %>% relocate(date, race_id)
```


```{r}
b %>%
  filter(
    race_id == 3
    ) %>% 
  arrange(result, won) %>% 
  relocate(race_id, result, won, contains('horse'))

```

# EDA

## Factor vars

```{r}
a %>% inspect_cat()
```

```{r}
 a %>% count(venue, config, distance, going, horse_ratings, sort = T) %>% mutate(pct = n/sum(n))
```

```{r}
a %>% freq
#group together 'going' levels less than 1% into an 'Other' category, or filter them out altogether
#ditto for horse_ratings
```
### important filters
```{r}
(top9.horse.ratings = a %>% count(horse_ratings, sort = T) %>% mutate(pct = n/sum(n)) %>% filter(pct > 0.02))
(top3.going = a %>% count(going, sort = T) %>% mutate(pct = n/sum(n)) %>% filter(pct > 0.02))

a = a %>%
  filter(
    horse_ratings %in% top9.horse.ratings$horse_ratings,
    going %in% top3.going$going
  )

a %>% freq
```

### filtering by venue
```{r}
a.st = a %>% filter(venue == 'ST') %>% select(-date)
a.hv = a %>% filter(venue == 'HV') %>% select(-date)
```

```{r}
(a.st.grouped = a.st %>% 
  group_by(config, distance, going, horse_ratings) %>% 
  summarise(across(
    .cols = contains('sec_'),
    .fns =  list(mean = mean, median = median)
  )) %>%
  left_join(a.st %>% count(config, distance, going, horse_ratings, sort = T) %>% mutate(pct = n/sum(n))) %>% 
  arrange(-pct) %>% 
  relocate(n, pct))
```

```{r}
(a.hv.grouped = a.hv %>% 
  group_by(config, distance, going, horse_ratings) %>% 
  summarise(across(
    .cols = contains('sec_'),
    .fns =  list(mean = mean, median = median)
  )) %>%
  left_join(a.st %>% count(config, distance, going, horse_ratings, sort = T) %>% mutate(pct = n/sum(n))) %>% 
  arrange(-pct) %>% 
  relocate(n, pct))
```

